\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-01-30 paper/main.tex
minor updates

\subsection{Choosing a cross-validation splitting scheme}
As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked. Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties. We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset. \\
Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}. (1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}). This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance. (2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others. In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors. (3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those. Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\

@@

This scenario is motivated by the need to identify drugs that can target a
specific protein that may be associated with a disease or abnormal
phenotypes.
Our aim is further to combine two different types of features:
structural information based on protein structure and molecular drug composition, and their respective observable characteristics. \\

>>>>

\subsection{Choosing a cross-validation splitting scheme}

% As in machine learning inference is derived from the underlying data,
% models and data are naturally and intrinsically linked. Thus, the more
% we understand about the pitfalls and biases within the data, the more
% we can try to bypass these difficulties. We will hereby abbreviate,
% e.g., a cross validation splitting scheme as \glqq split\grqq{}
% determining the train, validation and test subset of a given dataset.

Within DTI prediction, one can observe basic biases resulting from the
underlying datasets \citep{Pahikkala2014}. First, novel drugs are
often designed by altering non-functional components of a drug,
leading to two and more very similar drugs designed to target the same
proteins \cite{Overington2006}. This can lead to a bias when it leads
to \glqq hidden duplicates\grqq{} that can distribute among the
train--test split, resulting in a better (measured) predictive
performance than would be expected when the model is applied to
genuinely unknown interactions. Second, some proteins (which we will
denote as \textit{hub proteins}) have significantly more known
interactions with drugs than others. In the STITCH database, $5\%$
of the proteins have $40\%$ of the interactions, and similar
distributions are present in the Yamanishi and Drugbank datasets
\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
these proteins can often increase predictive performance while again
not reflecting the actual performance when applied to a new protein
(e.g., a protein for which no interactions are known). This bias in
the number of drugs targeting certain proteins may be the result of
study bias where more \glqq valuable\grqq{} proteins have more drugs
designed to target them due to their involvement in more common
diseases (or diseases for which drugs can be more profitably
marketed).

The aim of \name is to predict candidate drugs that target a given
protein; the challenge is to develop a training and evaluation scheme
that does not simply overfit to the inherent biases in training and
testing data.

@@

This scenario is motivated by the need to identify drugs that can target a
specific protein that may be associated with a disease or abnormal
phenotypes.
Our aim is further to combine two different types of features:
structural information based on protein structure and molecular drug composition, and their respective observable characteristics.
